---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eaflood

<!-- badges: start -->
[![R-CMD-check](https://github.com/DavidASmith/eaflood/workflows/R-CMD-check/badge.svg)](https://github.com/DavidASmith/eaflood/actions)
<!-- badges: end -->

The `eaflood` package provides access to data from the 
[Environment Agency FloodMonitoring API](https://environment.data.gov.uk/flood-monitoring/doc/reference) 
covering flood areas in England.

The following data are available:

- Flood warnings
- Flood areas
- Monitoring stations
- Measures (_e.g._ level and flow recorded at the monitoring stations)
- Readings values for each measure over the previous 4 weeks.

## Installation

You can install the development version of eaflood from 
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DavidASmith/eaflood")
```

## Examples

Here a few examples of what you can do with `eaflood`:

```{r package}
library(eaflood)
```

Get current flood warnings:

```{r flood_warnings, message=FALSE, warning=FALSE}
plot_flood_warnings(interactive = FALSE)
```
Get flood areas for a given location...

```{r flood_areas, message=FALSE, warning=FALSE}
plot_flood_areas(lat = 53.38,
                 lon = -1.46,
                 dist = 10, 
                 interactive = FALSE)
```
Get details of available monitoring stations...

```{r stations, message=FALSE, warning=FALSE}
plot_stations(town = "Gloucester", interactive = FALSE)
```
Get recent readings for a particular measure...

```{r plot_readings}
plot_measure_readings("L0607-level-stage-i-15_min-m")
```


See the vignette for more information.
