---
title: "Introduction to eaflood"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to eaflood}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(eaflood)
library(sf)
library(tmap)
library(tibble)
library(ggplot2)
library(lubridate)
library(dplyr)
```

The `eaflood` package provides access to data from the 
[Environment Agency FloodMonitoring API](https://environment.data.gov.uk/flood-monitoring/doc/reference) 
covering flood areas in England.

The package also includes functions for plotting the returned data. This 
vignette aims to help you get started using the package to explore flood 
warnings and river data.

# Get flood warnings

The Environment Agency issue warnings of floods that cover specific warning or 
alert areas. The floods API provides a listing of all current such warnings and 
is updated every 15 minutes. Flood warnings can be retrieved with 
`get_flood_warnings()`.

```{r}
flood_warnings <- get_flood_warnings()

if(!is.null(flood_warnings)) {
  tibble(flood_warnings)
}
```

Note that you don't need to use `tibble()` as above. We're only doing that here 
to print the contents of the dataframe to the vignette in a more easily 
readable format.

You can provide arguments to `get_flood_warnings()` to limit returned warnings 
to a particular county or within a given distance of a particular location. 
See ?get_flood_warnings for more information.

You can also visualise flood warnings in an interactive map.

```{r fig.height=4, fig.width=6}
plot_flood_warnings()
```

Note that the flood warnings shown here are as current when this vignette was 
generated, not those currently active. If no flood warnings were available when 
the vignette was built, the map will not be displayed.

# Flood areas

The flood areas API provide information on the geographic regions to which a 
given flood alert or warning may apply. These comprise Flood Alert Areas and 
Flood Warning Areas. A Flood Alert Area is a geographical area where it is 
possible for flooding to occur from rivers, sea and in some locations, 
groundwater. A single Flood Alert Area may cover a large portion of the 
floodplain, may contain multiple river catchments of similar characteristics 
and may contain a number of Flood Warning Areas. A Flood Warning Area is a 
geographical area where Environment Agency expect flooding to occur and which 
they provide a Flood Warning Service.

Full information on the areas is available from the Environment Agency Spatial 
Data Catalogue as a downloadable file and via a Web Feature Service or Web 
Mapping Service. For convenience we here provide the feature information for 
each area as a simple JSON format, including a specification of the polygon for 
each area (as a geoJSON feature in WGS84 coordinates). A typical application 
should maintain a local copy of the geographic information rather that rely on 
on-demand downloads of the rather large polygon files.

Each flood warning provides a link (floodArea) to the URI of the flood area to 
which it applies.

Use `get_flood_areas()` to return the flood areas from the API. For example, 
to return all flood areas within 10km of the given coordinates... 

```{r}
flood_areas <- get_flood_areas(lat = 53.38,
                               lon = -1.46,
                               dist = 10)
tibble(flood_areas)
```

You can also limit the number of returned items using the `limit` argument...

```{r}
flood_areas <- get_flood_areas(lat = 53.38,
                               lon = -1.46,
                               dist = 10, 
                               limit = 5)
tibble(flood_areas)
```

Additionally, you can generate an interactive plot of the flood areas...

```{r fig.height=4, fig.width=6}
plot_flood_areas(lat = 53.38,
                 lon = -1.46,
                 dist = 10)
```

Note that any invalid geometries (as evaluated by `sf::st_is_valid()`) present 
in the GeoJSON polygons returned by the API are not plotted, but are indicated 
in a warning message.

# Get stations

The real-time data API provides information on readings of water levels and 
flows taken at a variety of measurement stations. A given station may provide 
more than one set of measurements, for example, both water level and flow rate 
or water level at two different spots (e.g. up-stream and down-stream of a 
sluice or weir). The API provides metadata on these stations and on the 
different measures available from each one, as well as the readings themselves.

Use `get_stations()` to get details of measurement stations from the API. For 
example, to return all stations having a `label` containing `"Dronfield"`... 

```{r}
stations <- get_stations(search = "Dronfield")
tibble(stations)
```

Or to return all stations having a `town` of `"Leeds"`.

```{r}
stations <- get_stations(town = "Leeds")
tibble(stations)
```

# Get measures

In additional to listing stations, which includes the information on the 
measures available from each station, you can list the available measures 
directly with `get_measures()`

For example, to return all measures available via the API...

```{r}
measures <- get_measures()
tibble(measures)
```

To return only those measures which record river flow...

```{r}
measures <- get_measures(parameter = "flow")
tibble(measures)
```

To return only those measures for a particular station, pass the station 
reference to the `station_reference` argument. Station reference is given in 
the `notation` column of the dataframe returned by `get_stations`. 

```{r}
measures <- get_measures(station_reference = "4169TH")
tibble(measures)
```

# Get readings

Readings for each of the published measures are available for the recent period 
(up to the last four weeks). Each reading comprises a reference to the measure 
being read (identified by its URI), a date time stamp for when the reading 
applies and a numeric value. The set of readings is updated every 15 minutes, 
but the individual measures may be updated less frequently than this.

To return all readings, use `get_readings()`.

```{r}
readings <- get_readings(sorted = FALSE, limit = 10)
tibble(readings)
```

Note that this is unlikely to be useful unless other arguments are used to 
filter the returned data. Also note that using `sorted = TRUE` (the default) 
with an unfiltered query may cause the operation to time out.

See `?get_readings` for a full description of the filtering available. As 
examples, use the following to return all latest readings for a particular 
station...

```{r}
readings <- get_readings(station = "3210TH", latest = TRUE)
tibble(readings)
```

Or the following to get all readings from today for a particular measure...

```{r}
readings <- get_readings(measure = "L0607-level-stage-i-15_min-m", 
                         today = TRUE)
```

Or the following to return all readings from the last three days for a 
particular measure...

```{r}
readings <- get_readings(measure = "L0607-level-stage-i-15_min-m", 
                                 since = today() - 3)

```

We can plot those readings like this...

```{r fig.height=4, fig.width=6}
readings %>% 
  ggplot(aes(dateTime, value)) + 
  geom_line() + 
  labs(title = "River level for measure L0607-level-stage-i-15_min-m") +
  xlab("Date") +
  ylab("Level (m)")
```


```{r fig.height=4, fig.width=6}
plot_measure_readings("L0607-level-stage-i-15_min-m")
```

